#Reproducible Research: Peer Assessment 1  
  
  
##Loading and preprocessing the data
```{r loaddata}
#Unzip the csv
unzip("repdata-data-activity.zip")

#read csv
activity <- read.csv("activity.csv")
```
  
  
##What is mean total number of steps taken per day?
#### 1. Calculate the total number of steps taken per day
```{r}
totalStepPerDay <- tapply(activity$steps, activity$date, sum, na.rm = TRUE)
```

#### 2. Make a histogram of the total number of steps taken each day
```{r}
library(ggplot2)
qplot(totalStepPerDay, xlab = "Total steps per day")
```

#### 3. Calculate and report the mean and median of the total number of steps taken per day
```{r}
mean(totalStepPerDay)
median(totalStepPerDay)
```

##What is the average daily activity pattern?

#### 1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
library(ggplot2)
averageStepsPerInterval <- aggregate(x = list(steps = activity$steps), by = list(interval = activity$interval), FUN = mean, na.rm = TRUE)
ggplot(data = averageStepsPerInterval, aes(x = averageStepsPerInterval$interval, y = averageStepsPerInterval$steps)) + geom_line() + xlab("5-minute interval") + ylab("Average steps taken")
```

#### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
averageStepsPerInterval[which.max(averageStepsPerInterval$steps),]
```

##Imputing missing values
#### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
missingValues <- is.na(activity$steps)
table(missingValues)
```

#### 2. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
filledData <- activity

for(index in 1:length(filledData[[1]])) {
    if(is.na(filledData[index,]$steps)) {
        filledData[index,]$steps <- averageStepsPerInterval[averageStepsPerInterval$interval == filledData[index,]$interval, "steps"]
    }
}
```

#### 3. Make a histogram of the total number of steps taken each day...
```{r}
totalStepPerDayAfterFillingData <- tapply(filledData$steps, filledData$date, sum, na.rm = TRUE)
qplot(totalStepPerDayAfterFillingData, xlab = "Total steps per day")
```

#### ...Calculate and report the mean and median total number of steps taken per day.
```{r}
mean(totalStepPerDayAfterFillingData)
median(totalStepPerDayAfterFillingData)
```

##Are there differences in activity patterns between weekdays and weekends?
#### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
weekday <- function(date) {
    day <- weekdays(date)
    if(day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday"))
    {
        return("Weekday")
    } else if (day %in% c("Saturday", "Sunday")) {
        return("Weekend")
    }
    return("Invalid date")
}
filledData$day <- sapply(as.Date(filledData$date), FUN = weekday)
```

#### 2. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis).
```{r}
averageStepsPerWeekday <- aggregate(steps ~ interval + day, data = filledData, mean)
ggplot(averageStepsPerWeekday, aes(interval, steps)) + geom_line() + facet_grid(day ~ .) + xlab("5 minute interval") + ylab("Number of steps")
```